---
title: "GLM"
date: "2023-03-09"

  output: github_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE)
```



```{r }
library(tidyverse)
library(moderndive)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)
```


#Binary logistic regression with one numerical explanatory variable

```{r }
evals.gender <- evals %>%
                  select(gender, age)
evals.gender
```


look at a boxplot of by to get an initial impression of the data

```{r}
ggplot(data = evals.gender, aes(x = gender, y = age, fill = gender)) +
  geom_boxplot() +
  labs(x = "Gender", y = "Age")+ 
  theme(legend.position = "none")

```

## Log-odds
The logistic regression model with **gender** as the response and **age** as the explanatory variable is given by:

```{r}
model <- glm(gender ~ age, data = evals.gender, 
             family = binomial(link = "logit"))
model
model %>%
  summary()
#or
summ(model)
```


*Baseline* in R being taken as the one which comes first alphabetically, in this model is **female**.
```{r}
levels(evals.gender$gender)
```

extract the estimated coefficients:
```{r}
mod1coefs <- round(coef(model), 2)
mod1coefs
```


\begin{align}

\ln\left(\frac{p}{1-p}\right) &= \alpha + \beta \cdot
\textrm{age} = `r mod1coefs[1]` + `r mod1coefs[2]` \cdot \textrm{age} \nonumber

\end{align}

The *log-odds* of the instructor being male increase by 0.06 for every one unit increase in.
a 95% confidence interval for these log-odds:

```{r}
library(kableExtra)
confint(model) %>%
  kable()
```

#
```{r}
mod.coef.logodds <- model %>%
                      summary() %>%
                      coef()
mod.coef.logodds

#logodds.lower
age.logodds.lower <- mod.coef.logodds["age", "Estimate"] - 
                      1.96 * mod.coef.logodds["age", "Std. Error"]
age.logodds.lower

#logodds.upper
age.logodds.upper <- mod.coef.logodds["age", "Estimate"] + 
                      1.96 * mod.coef.logodds["age", "Std. Error"]
age.logodds.upper
```

Hence the point estimate for the log-odds is 0.06, which has a corresponding 95% confidence interval of (0.04, 0.08),This can be displayed graphically :

```{r}
#The log-odds of age for male instructors.
plot_model(model, show.values = TRUE, transform = NULL,
           title = "Log-Odds (Male instructor)", show.p = FALSE)
```

 let's add the estimates of the log-odds to our data set:

```{r}
evals.gender <- evals.gender %>%
                  mutate(logodds.male = predict(model))
evals.gender
evals.gender[evals.gender$age==29 ,3]
```




## odds
odds scale as it is easier to interpret an odds-ratio as opposed to the log-odds-ratio.

```{r}
model %>%
 coef() %>%
  exp()
```

```{r}
age.odds.lower <- exp(age.logodds.lower)
age.odds.upper <- exp(age.logodds.upper)
age.odds.lower
age.odds.upper
```
Hence the point estimate for the odds is 1.06, which has a corresponding 95% confidence interval of (1.04, 1.09).
displayed graphically :
```{r}
plot_model(model, show.values = TRUE, axis.lim = c(1,1.5),
           title = "Odds (Male instructor)", show.p = FALSE)
```
```{r}
evals.gender <- evals.gender %>%
                  mutate(odds.male = exp(logodds.male))
evals.gender

evals.gender[evals.gender$age==56 ,4]
```



## Probabilities
```{r}
p.num <- exp(mod.coef.logodds["(Intercept)", "Estimate"] + mod.coef.logodds["age", "Estimate"] * 52)
p.denom <- 1 + p.num
p.num / p.denom
```
or use **plogis()** function

```{r}
plogis(mod.coef.logodds["(Intercept)", "Estimate"] + mod.coef.logodds["age", "Estimate"] * 52)
```

add the probabilities to our data:
```{r}
evals.gender <- evals.gender %>%
                  mutate(probs.male = fitted(model))
evals.gender

#estimated probabilities.
predict(model, type = "response")

#??????????
evals.gender[evals.gender$age==33 & evals.gender$gender=="female",5]

#??????????

```


plot the probability:
```{r}
ggplot(data = evals.gender, aes(x = age, y = probs.male)) +
  geom_smooth(method="glm", 
              method.args = list(family="binomial"), 
              se = FALSE) +
  labs(x = "Age", y = "Probability of instructor being male")
```

```{r estimated probabilities}
plot_model(model, type = "pred", title = "",
            axis.title = c("Age", "Prob. of instructor being male"))
```






#Binary logistic regression with one categorical explanatory variable

the binary categorical variable **ethnicity** as our explanatory variable
**gender** as the response
```{r}
evals.ethnic <- evals %>%
                  select(gender, ethnicity)
evals.ethnic
```

# summarise this data in a table format:
```{r summarise}
library(janitor)
evals %>% 
  tabyl(ethnicity, gender) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns() # To show original counts
```


visualize the distribution using a barplot:
```{r  a barplot}
ggplot(evals, aes(x= gender,  y = ..prop.., group=ethnicity, fill=ethnicity)) + 
    geom_bar(position="dodge", stat="count") +
    labs(y = "Proportion")
```


Now we shall fit a logistic regression model to determine whether the gender of a teaching instructor can be predicted from their ethnicity.


## Log-odds
logistic regression model:

```{r logistic model}
model.ethnic <- glm(gender ~ ethnicity, data = evals.ethnic, family = binomial(link = "logit"))

model.ethnic %>%
  summary()
```

```{r baseline}
levels(evals.ethnic$ethnicity)
```

95% confidence interval for these log-odds
```{r log-odds CI}
confint(model.ethnic) %>%
  kable()
```
#or calculate the CI:
```{r}
mod.ethnic.coef.logodds <- model.ethnic %>%
                            summary() %>%
                            coef()
mod.ethnic.coef.logodds

ethnic.logodds.lower <- mod.ethnic.coef.logodds["ethnicitynot minority", "Estimate"] - 
                        1.96 * mod.ethnic.coef.logodds["ethnicitynot minority", "Std. Error"]

ethnic.logodds.upper <- mod.ethnic.coef.logodds["ethnicitynot minority", "Estimate"] + 
                        1.96 * mod.ethnic.coef.logodds["ethnicitynot minority", "Std. Error"]

ethnic.logodds.lower
ethnic.logodds.upper
```
Hence the point estimate for the log-odds is 0.66, which has a corresponding 95% confidence interval of (0.13, 1.2).


displayed graphically 

```{r}
plot_model(model.ethnic, show.values = TRUE, transform = NULL,
           title = "Log-Odds (Male instructor)", show.p = FALSE)
```
add the estimates of the log-odds to our data set:

```{r add log-odds}
evals.ethnic <- evals.ethnic %>%
                  mutate(logodds.male = predict(model.ethnic))
evals.ethnic
```

## Odds
```{r}
model.ethnic %>%
 coef() %>%
  exp()
```


```{r}
# the number of instructors in the minority
pmin <- evals.ethnic %>%
              filter(ethnicity == "minority") %>%
              summarize(n()) %>%
              pull()
pmin

# the number of male instructors in the minority
pmin.male <- evals.ethnic %>%
              filter(ethnicity == "minority", gender == "male") %>%
              summarize(n()) %>%
              pull()
pmin.male

# the proportion/probability of males in the minority
prob.min.male <- pmin.male / pmin
prob.min.male 

# the odds of an instructor being male given they are in the minority
odds.min.male <- prob.min.male / (1 - prob.min.male)
odds.min.male
```

95% confidence interval for the odds
```{r}
ethnic.odds.lower <- exp(ethnic.logodds.lower)
ethnic.odds.lower
ethnic.odds.upper <- exp(ethnic.logodds.upper)
ethnic.odds.upper
plot_model(model.ethnic, show.values = TRUE,
           title = "Odds (Male instructor)", show.p = FALSE)
```

add the estimates of the odds:
```{r}
evals.ethnic <- evals.ethnic %>%
                  mutate(odds.male = exp(logodds.male))
evals.ethnic 
```

## Probabilities

The probabilities of an instructor being male given they are in the minority and not minority groups are:

```{r}
plogis(mod.ethnic.coef.logodds["(Intercept)", "Estimate"])
plogis(mod.ethnic.coef.logodds["(Intercept)", "Estimate"] + 
         mod.ethnic.coef.logodds["ethnicitynot minority", "Estimate"]) 
```
add the probabilities to our data:
```{r}
evals.ethnic <- evals.ethnic %>%
                  mutate(probs.male = fitted(model.ethnic))
evals.ethnic
```

produce the estimated probabilities:
```{r}
plot_model(model.ethnic, type = "pred", title = "",
            axis.title = c("Ethnicity", "Prob. of instructor being male"))
```




# Further tasks











